RProtoBuf: Efficient Cross-Language Data Serialization in R

### 2014-01-04 papers/rjournal/eddelbuettel-stokely.Rnw
another incomplete round of edits
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@703 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

Before we can create a new Protocol Buffer Message or parse a
serialized stream of bytes as a Message, we must read in the message
type specification from a \texttt{.proto} file.

@@

<<>>=
ls( "RProtoBuf:DescriptorPool" )
@

@@

<<>>=
p1 <- new( tutorial.Person )
p <- new( tutorial.Person, name = "Romain", id = 1 )
@

@@

<<>>=
p$name
p$id
p$email <- "francoisromain@free.fr"
@

@@

<<>>=
p[["name"]] <- "Romain Francois"
p[[ 2 ]] <- 3
p[[ "email" ]]
@

@@

<<>>=
writeLines( as.character( p ) )
@

@@

<<>>=
serialize( p, NULL )
@

@@

<<>>=
tf1 <- tempfile()
serialize( p, tf1 )
readBin( tf1, raw(0), 500 )
@

@@

<<>>=
tf2 <- tempfile()
con <- file( tf2, open = "wb" )
serialize( p, con )
close( con )
readBin( tf2, raw(0), 500 )
@

@@

<<>>=
# serialize to a file
p$serialize( tf1 )
# serialize to a binary connection
con <- file( tf2, open = "wb" )
p$serialize( con )
close( con )
@

@@

<<>>=
msg <- read( tutorial.Person, tf1 )
writeLines( as.character( msg ) )
@

@@

<<>>=
con <- file( tf2, open = "rb" )
message <- read( tutorial.Person, con )
close( con )
writeLines( as.character( message ) )
@

@@

<<>>=
# reading the raw vector payload of the message
payload <- readBin( tf1, raw(0), 5000 )
message <- read( tutorial.Person, payload )
@

@@

<<>>=
# reading from a file
message <- tutorial.Person$read( tf1 )
# reading from a binary connection
con <- file( tf2, open = "rb" )
message <- tutorial.Person$read( con )
close( con )
# read from the payload
message <- tutorial.Person$read( payload )
@

@@

Each R object stores an external pointer to an object managed by
the \texttt{protobuf} C++ library.
The \CRANpkg{Rcpp} package \citep{eddelbuettel2011rcpp} is used to
facilitate the integration of the R and C++ code for these objects.

@@

There are over 100 C++ functions that provide the glue code between
the member functions of the 6 primary Message and Descriptor classes
in the protobuf library.  Wrapping each method individually allows us
to add user friendly custom error handling, type coercion, and
performance improvements at the cost of a more verbose
implementation.  The RProtoBuf implementation in many ways motivated
the development of Rcpp Modules \citep{eddelbuettel2010exposing},
which provide a more concise way of wrapping C++ functions and classes
in a single entity.

The \texttt{RProtoBuf} package combines the \emph{R typical} dispatch
of the form \verb|method( object, arguments)| and the more traditional
object oriented notation \verb|object$method(arguments)|.
Additionally, \texttt{RProtoBuf} implements the \texttt{.DollarNames} S3 generic function
(defined in the \texttt{utils} package) for all classes to enable tab
completion.  Completion possibilities include pseudo method names for all
classes, plus dynamic dispatch on names or types specific to a given object.

>>>>

Before one can create a new Protocol Buffer Message or parse a
serialized stream of bytes as a Message, one must first read in the message
type specification from a \texttt{.proto} file.

@@

<<>>=
ls("RProtoBuf:DescriptorPool")
@

@@

<<>>=
p1 <- new(tutorial.Person)
p <- new(tutorial.Person, name = "Murray", id = 1)
@

@@

<<>>=
p$name
p$id
p$email <- "murray@stokely.org"
@

@@

<<>>=
p[["name"]] <- "Murray Stokely"
p[[ 2 ]] <- 3
p[[ "email" ]]
@

@@

<<>>=
writeLines(as.character(p))
@

@@

<<>>=
serialize(p, NULL)
@

@@

<<>>=
tf1 <- tempfile()
serialize(p, tf1)
readBin(tf1, raw(0), 500)
@

@@

<<>>=
tf2 <- tempfile()
con <- file(tf2, open = "wb")
serialize(p, con)
close(con)
readBin(tf2, raw(0), 500)
@

@@

<<>>=
# serialize to a file
p$serialize(tf1)
# serialize to a binary connection
con <- file(tf2, open = "wb")
p$serialize(con)
close(con)
@

@@

<<>>=
msg <- read(tutorial.Person, tf1)
writeLines(as.character(msg))
@

@@

<<>>=
con <- file(tf2, open = "rb")
message <- read(tutorial.Person, con)
close(con)
writeLines(as.character(message))
@

@@

<<>>=
# reading the raw vector payload of the message
payload <- readBin(tf1, raw(0), 5000)
message <- read(tutorial.Person, payload)
@

@@

<<>>=
# reading from a file
message <- tutorial.Person$read(tf1)
# reading from a binary connection
con <- file(tf2, open = "rb")
message <- tutorial.Person$read(con)
close(con)
# read from the payload
message <- tutorial.Person$read(payload)
@

@@

Each R object stores an external pointer to an object managed by
the \texttt{protobuf} C++ library.
The \CRANpkg{Rcpp} package \citep{eddelbuettel2011rcpp,eddelbuettel2013seamless} is used to
facilitate the integration of the R and C++ code for these objects.

@@

There are over 100 C++ functions that provide the glue code between
the member functions of the 6 primary Message and Descriptor classes
in the protobuf library.  Wrapping each method individually allows us
to add user friendly custom error handling, type coercion, and
performance improvements at the cost of a more verbose
implementation.  The RProtoBuf implementation in many ways motivated
the development of Rcpp Modules \citep{eddelbuettel2013exposing},
which provide a more concise way of wrapping C++ functions and classes
in a single entity.

The \texttt{RProtoBuf} package combines the \emph{R typical} dispatch
of the form \verb|method(object, arguments)| and the more traditional
object oriented notation \verb|object$method(arguments)|.
Additionally, \texttt{RProtoBuf} implements the \texttt{.DollarNames} S3 generic function
(defined in the \texttt{utils} package) for all classes to enable tab
completion.  Completion possibilities include pseudo method names for all
classes, plus dynamic dispatch on names or types specific to a given object.
